---
title: "Lipid analysis of sediment samples from Prasat"
author: "Liying Wang"
date: "7/10/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE)
```

# Introduction 

This is a document of data visualization for the preliminary lipid analysis of sediment samples

```{r read-data, message = FALSE}
library(tidyverse)
# from LiYingSed2020_GCMS_TIC.CSV, before methyl, just silylation
df <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTI_5zJpN_dywxtdFBKtON89jKXTzJencesAWjtO9yDZ4_LljkiLu9_EkP9XeIF8VrQxS2H_iU1ZnC4/pub?output=csv")
ixd <- which(str_detect(df$Path, "D"))
dfs <- split(df, cumsum(1:nrow(df) %in% ixd))
dfs_data <- dfs[c(FALSE, TRUE)]
dfs_name <- dfs[c(TRUE, FALSE)]
names(dfs_data) <- map(dfs_data, ~.x$Path[1])
names(dfs_name) <- map(dfs_name, ~.x$Sample[1])
dfs <- map_df(dfs_data, ~.x[-1, ] %>% 
              mutate_all(as.numeric), 
              .id = 'sample')

```

```{r tidy-data}
# get sample number
dfs <- 
  dfs %>% 
  mutate(sample = str_extract(sample, pattern = "_00[0-9]")) %>% 
  mutate(Sample = case_when(sample == "_002" ~ substr(names(dfs_name)[1], start = 1, stop = 5),
                            sample == "_003" ~ substr(names(dfs_name)[2], start = 1, stop = 5),
                            sample == "_004" ~ substr(names(dfs_name)[3], start = 1, stop = 5),
                            sample == "_005" ~ substr(names(dfs_name)[4], start = 1, stop = 5),
                            sample == "_006" ~ substr(names(dfs_name)[5], start = 1, stop = 5),
                            sample == "_007" ~ substr(names(dfs_name)[6], start = 1, stop = 5),
                            sample == "_008" ~ substr(names(dfs_name)[7], start = 1, stop = 5),
                            sample == "_009" ~ substr(names(dfs_name)[8], start = 1, stop = 5)))
```

# Plot

Plot the GCMS results of all samples

```{r plot-all}
ggplot(dfs,
       aes(Path, File)) +
  geom_line() +
  labs(x = "Retention time",
       y = "Intensity",
       title = names(dfs[1])) +
  theme_minimal() +
  facet_wrap(~Sample)
```

Select and plot the samples that show interesting results of organic compounds over time

```{r plot-selected-sample}
# select interested layers
interest <- 
  dfs %>% 
  filter(Sample == c("PT045", "PT095", "PT145"))

highlight_df <- 
  interest %>% 
  filter(File >= 200000) %>% 
  mutate(fatty = round(Path)) %>% 
  group_by(Sample, fatty) %>% 
  arrange(desc(File)) %>% 
  distinct(fatty, .keep_all = TRUE) %>% 
  filter (!fatty %in% c(49, 29, 22, 8)) %>% 
  mutate(fatty_label = case_when(fatty == 28 ~ "C16:0", #Palmatic
                                 fatty == 31 ~ "C18:0", #Stearic
                                 #fatty == 22 ~ "Myristic",
                                 fatty == 36 ~ "C22:0", #Behenic
                                 fatty == 46 ~ "C28:0", #Montanic
                                 fatty == 48 ~ "C30:0")) #Melissic

ggplot(interest,
       aes(Path, File)) +
  geom_line() +
  geom_point(data = highlight_df, aes(Path, File), color="red", size=1) +
  ggrepel::geom_text_repel(data = highlight_df, 
                           aes(label = fatty_label),
                           size = 2.5,
                           vjust = -1.2, 
                           hjust = 0.4,
                           show.legend = FALSE) +
  labs(x = "Retention time",
       y = "Intensity") +
       #title = names(dfs[1]))
  theme_minimal() +
  facet_wrap(~Sample, ncol = 3) +
  theme(strip.text.x = element_text(vjust = 3,
                                    hjust = 0,
                                    size = 6))
```

# Results

GC-MS results show higher proportion of long-chain plant-wax fatty acids in the early period (PT145). In the middle period (PT095), an increasing proportion of animal fats were identified, and they became dominant fatty acids in the late period (PT045)
